---
title: "EC2880C Group Project Data Analysis"
author: "Adam Cheng"
bibliography: https://raw.githubusercontent.com/adamcheng7/EC2880C-Group-Project/refs/heads/main/syntax/data_analysis/bibliography/data_analysis_citations.bib
csl: https://raw.githubusercontent.com/adamcheng7/EC2880C-Group-Project/refs/heads/main/syntax/data_analysis/bibliography/apa_6th_edition.csl
format:
    html:
      theme: cosmo
      fontsize: 14px
      page-layout: article
      html-math-method: mathml
      code-overflow: wrap
      grid:
        sidebar-width: 250px
        body-width: 950px
        margin-width: 250px
        gutter-width: 1.5rem
      code-links:
      - text: Project Repository
        icon: github
        href: https://github.com/adamcheng7/EC2880C-Group-Project
      number-sections: true
      toc: true
      toc-expand: true
      toc-location: left
      df-print: paged
      code-fold: false
      self-contained: true
execute:
  echo: true
  warning: false
  message: false
editor: visual
---

# Introduction

This document aims to produce a line plot graphic comparing the annual export growth rate of Argentina to the annual average $E_{ARS/USD}$ (ARS=X) exchange rate between 1990-2005.

# Loading R Packages

```{r}
#| label: loading_packages
#| output: false

# List of necessary packages
packages <- c("rlang",
              "tidyverse",
              "repr",
              "janitor",
              "haven",
              "patchwork"
              )

# Install packages that are not installed
new_packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new_packages) > 0) install.packages(new_packages)

# Loading the necessary R packages
library(rlang)
library(tidyverse)
library(repr)
library(janitor)
library(haven)
library(patchwork)

# Setting seed for reproducibility
set.seed(2025)

# Disabling scientific notations
options(scipen = 999)
```

# Reading Datasets

```{r}
#| label: reading_data

# Reading World Bank data
export_data_url <- "https://raw.githubusercontent.com/adamcheng7/EC2880C-Group-Project/refs/heads/main/data/arg_export_wb.csv"
export_data <- read_csv(export_data_url) |> slice(1)

export_data

# Reading IMF data
exchange_data_url <- "https://raw.githubusercontent.com/adamcheng7/EC2880C-Group-Project/refs/heads/main/data/arg_ars-usd_imf.csv"
exchange_data <- read_csv(exchange_data_url)

exchange_data
```

# Wrangling and Cleaning Datasets

## Wrangling and Cleaning International Monetary Fund (IMF) Data

Source: @imf

```{r}
#| label: cleaning_imf

# Renaming and selecting necessary columns
exchange_wrangled <- exchange_data |>
  rename("country_name"="COUNTRY",
         "year"="TIME_PERIOD",
         "ars_usd"="OBS_VALUE") |>
  select(country_name,year,ars_usd)

# Filtering year range to 1990-2005
exchange_cleaned <- exchange_wrangled |>
  filter(year >= 1990 & year <= 2005)

exchange_cleaned
```

## Wrangling and Cleaning World Bank (WB) Data

Source: @wb

```{r}
#| label: cleaning_wb

export_wrangled <- export_data |>
  clean_names() |>
  pivot_longer(cols = x1960_yr1960:x2024_yr2024, names_to = "year", values_to = "value") |> # Pivot dataset to long format
  mutate(year = as.numeric(str_extract(year, "\\d{4}"))) |> # Extracts the 4-digit year number from the original dataset's incompatible year format
  select(country_name, year, value)|>
  rename("export_growth"="value")

# Filtering year range to 1990-2005
export_cleaned <- export_wrangled |>
  filter(year >= 1990 & year <= 2005)

export_cleaned
```

# Merging Datasets

```{r}
#| label: merging_data

# Joining datasets by year and country
complete_dataset <- full_join(export_cleaned, exchange_cleaned,
                              by = c("country_name","year"))

complete_dataset
```

# Producing Graphics

## Argentina's Annual $E_{ARS/USD}$ Average vs. Annual Export Growth Rate (Line Plot)

```{r}
#| label: fig-export_exchange
#| fig-width: 8
#| fig-height: 9
#| fig-cap: "Argentina's annual export growth rate and ARS/USD exchange rate between 1990 - 2005"

# Producing line plot

## Export plot (top)
export_plot <- complete_dataset |>
  ggplot(aes(x = year, y = export_growth)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = 1992, linetype = "dashed", color = "#004488") +
  geom_vline(xintercept = 1995, linetype = "dashed", color = "#DDAA33") +
  geom_vline(xintercept = 2002, linetype = "dashed", color = "#BB5566") +
  scale_x_continuous(breaks = seq(1990, 2006, by = 2)) +
  labs(y = "Export Growth Rate (Annual %)", x = "Year", title = "Argentina's annual export growth rate and ARS/USD exchange rate between 1990 - 2005")

## Exchange rate plot (bottom)
exchange_plot <- complete_dataset |>
  ggplot(aes(x = year, y = ars_usd)) +
  geom_line() +
  geom_point() +
  geom_vline(aes(xintercept = 1992, color = factor("Convertibility Plan", 
                                                   levels = c("Convertibility Plan", 
                                                              "Mexican Peso Devaluation", 
                                                              "Convertibility Plan Abandoned"))), 
             linetype = "dashed") +
  geom_vline(aes(xintercept = 1995, color = factor("Mexican Peso Devaluation", 
                                                   levels = c("Convertibility Plan", 
                                                              "Mexican Peso Devaluation", 
                                                              "Convertibility Plan Abandoned"))), 
             linetype = "dashed") +
  geom_vline(aes(xintercept = 2002, color = factor("Convertibility Plan Abandoned", 
                                                   levels = c("Convertibility Plan", 
                                                              "Mexican Peso Devaluation", 
                                                              "Convertibility Plan Abandoned"))), 
             linetype = "dashed") +
  scale_color_manual(name = "Events", 
                     values = c("Convertibility Plan" = "#004488",
                                "Mexican Peso Devaluation" = "#DDAA33",
                                "Convertibility Plan Abandoned" = "#BB5566")) +
  scale_x_continuous(breaks = seq(1990, 2006, by = 2)) +
  theme(legend.position = "bottom") +
  labs(y = "ARS per unit of USD", x = "Year")

## Combining plots
combined_plot <- export_plot / exchange_plot

combined_plot

# Uncomment to save plot as a png

# ggsave("arg_export_exchange_plot.png", plot = combined_plot)
```
